---
title: "CEA anticoag - base case"
author: "Emily O'Neill, Andrew Huang"
date: "2025-04-13"
output: html_document
---

```{r} 
##remove any variables in R's Memory 
rm(list = ls()) 

##Install packages for Markov Model
library(here)
library(tidyverse)

# Load functions
source(here::here("functions", "function_create-sa.R"))
source(here::here("functions", "function_make-psa-obj.R"))
source(here::here("functions", "function_summary-psa.R"))
source(here::here("functions", "function_plot-psa.R"))

``` 

Read-in base case results for riv and war

```{r}

sim_results_war_noMC <- readRDS(here::here("output","results_5yr_war_noMC.rds"))
sim_results_riv_noMC <- readRDS(here::here("output","results_5yr_riv_noMC.rds"))

```

Calculate ICER for base case

```{r} 
###CEA 

##Store the estimated cost of each medication 
v.c <- c(sim_results_war_noMC$tc_hat, 
         sim_results_riv_noMC$tc_hat)
##store the estimated QALY of each medication  
v.e <- c(sim_results_war_noMC$te_hat, 
         sim_results_riv_noMC$te_hat)


## ICER ##
delta.c <- v.c[2] - v.c[1]            # calculate incremental costs between rivaroxaban and warfarin 
delta.e <- v.e[2] - v.e[1]            # calculate incremental QALYs between rivaroxaban and warfarin
ICER <- delta.c / delta.e             # calculate the ICER
results <- list("delta.c" = delta.c, 
                "delta.e" = delta.e, 
                "ICER" = ICER)  # store the values in a new variable
saveRDS(results, here::here("output","results_5yr_basecase.rds"))

# Create full incremental cost-effectiveness analysis table
table_markov <- data.frame(
  round(v.c, 0),              # costs per arm
  round(v.e, 4),              # health outcomes per arm
  c("", round(delta.c, 0)),   # incremental costs
  c("", round(delta.e, 4)),   # incremental QALYs
  c("", round(ICER, 0))       # ICER
)
rownames(table_markov) = c("Warfarin", "Rivaroxaban")  # name the rows
colnames(table_markov) = c("Costs", "QALYs","Incremental Costs", "QALYs Gained", "ICER") # name the columns
table_markov                    # print the table 
```

Read-in MC simulation results for riv and war

```{r}

sim_results_war <- readRDS(here::here("output","results_5yr_war.rds"))
sim_results_riv <- readRDS(here::here("output","results_5yr_riv.rds"))

sim_results_war_param <- sim_results_war %>% select(v.cost, v.util, pr.ne_ne, pr.ne_vr, pr.ne_mb, pr.ne_nmb, pr.mb_ot, pr.nmb_ot)
sim_results_riv_param <- sim_results_riv %>% select(v.cost, v.util, pr.ne_ne, pr.ne_vr, pr.ne_mb, pr.ne_nmb, pr.mb_ot, pr.nmb_ot)

```

Calculate ICERs for MC simulations

```{r}

sim_results_cost <- data.frame(war = sim_results_war$tc_hat,
                               riv = sim_results_riv$tc_hat)
sim_results_effectiveness <- data.frame(war = sim_results_war$te_hat,
                                        riv = sim_results_riv$te_hat)

psa <- make_psa_obj(cost = sim_results_cost, effectiveness = sim_results_effectiveness, parameters = NULL,
                    strategies = c("Warfarin", "Rivaroxaban"), currency = "$", other_outcome = NULL)

# Save psa object
saveRDS(psa, here::here("output","results_5yr_psa.rds"))

# See mean ICER
summary.psa(psa, calc_sds=TRUE)

```

Distribution of ICER estimates

Table
```{r}
sim_results_ICER <- as.data.frame(
  cbind(inc_cost = sim_results_cost$riv - 
                   sim_results_cost$war,
        inc_effectiveness = sim_results_effectiveness$riv -                                           sim_results_effectiveness$war)) %>%
  mutate(ICER = inc_cost/inc_effectiveness,
         name = "MC")

mean(sim_results_ICER$ICER)
sd(sim_results_ICER$ICER)

library(boot)
set.seed(2025)
boot.icer = boot(data=sim_results_ICER$ICER,
                 R=1000,
                 statistic=function(x,id){ mean(x[id])})
boot.ci(boot.icer, type="bca")

```

Plot (ICERs)

```{r}
sim_results_ICER_boxplot <- sim_results_ICER %>%
  add_row(inc_cost = delta.c,
          inc_effectiveness = delta.e,
          ICER = ICER,
          name = "base")

ggplot(sim_results_ICER_boxplot, aes(x=name, y=ICER)) + 
  geom_boxplot() + 
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = c(-5000, 0)) +
  coord_flip() +
  xlab("Analysis") + 
  ylab("ICER ($/QALY)")

```

PSA plot (CERs)

```{r}
plot.psa(psa) +
  ggthemes::scale_color_colorblind() +
  ggthemes::scale_fill_colorblind() +
  xlab("Effectiveness (QALYs)") +
  guides(col = guide_legend(nrow = 2)) +
  theme(legend.position = "right")
```

Range of CERs (highest cost / lowest effectiveness)

```{r}
data.frame(
  Strategy = c("Warfarin", "Rivaroxaban"),
  max_cost = c(max(psa$cost$Warfarin), max(psa$cost$Rivaroxaban)),
  min_cost = c(min(psa$cost$Warfarin), min(psa$cost$Rivaroxaban)),
  max_effectiveness = c(max(psa$effectiveness$Warfarin),
                        max(psa$effectiveness$Rivaroxaban)),
  min_effectiveness = c(min(psa$effectiveness$Warfarin),
                        min(psa$effectiveness$Rivaroxaban)),
  best_CER = c(min(psa$cost$Warfarin)/max(psa$effectiveness$Warfarin),
               min(psa$cost$Rivaroxaban)/max(psa$effectiveness$Rivaroxaban)),
  worst_CER = c(max(psa$cost$Warfarin)/min(psa$effectiveness$Warfarin),
                max(psa$cost$Rivaroxaban)/min(psa$effectiveness$Rivaroxaban))
)
```

% of CER simulations where rivaroxaban is more cost-effective than warfarin

```{r}
# Simulated CERs for warfarin
psa_cer_war <- round(psa$cost$Warfarin / psa$effectiveness$Warfarin, digits=2)
# Simulated CERs for rivaroxaban
psa_cer_riv <- round(psa$cost$Rivaroxaban / psa$effectiveness$Rivaroxaban, digits=2)

# simulated CERs for rivaroxaban that are more cost effective than warfarin
psa_cer <- expand.grid(war = psa_cer_war, riv = psa_cer_riv) %>%
  distinct %>%
  mutate(riv_costeffective = ifelse(riv < war, 1, 0))

sum(psa_cer$riv_costeffective)/nrow(psa_cer)

```

% of CER simulations where rivaroxaban is more cost-effective than the best warfarin CER

```{r}
# Simulated CERs for warfarin
psa_cer_war <- round(psa$cost$Warfarin / psa$effectiveness$Warfarin, digits=2)
psa_cer_war <- sort(psa_cer_war)
# Simulated CERs for rivaroxaban
psa_cer_riv <- round(psa$cost$Rivaroxaban / psa$effectiveness$Rivaroxaban, digits=2)
psa_cer_riv <- sort(psa_cer_riv)

# simulated CERs for rivaroxaban that are more cost effective than best warfarin
# i.e., lower than the minimum CER for warfarin
psa_cer_riv_costeffective <- psa_cer_riv[psa_cer_riv < psa_cer_war[1]]
length(psa_cer_riv_costeffective) / length(psa_cer_riv)
```