---
title: "CEA anticoag - MC all"
author: "Emily O'Neill, Andrew Huang"
date: "2025-04-30"
output: html_document
---

MC draws for event rates, costs and utilities (1 year)

Setup

```{r} 
##Cost utility analysis of Rivaroxaban in the prevention of VTE reoccurrence versus standard or care with warfarin: A markov Model-Based Simulation  

##remove any variables in R's Memory 
rm(list = ls()) 

##Install packages for Markov Model 
library(here)
library(tidyverse)

# Load functions
source(here::here("functions", "function_markov.R"))
source(here::here("functions", "function_plot-trace.R"))
source(here::here("functions", "function_create-sa.R"))
source(here::here("functions", "function_make-psa-obj.R"))
source(here::here("functions", "function_summary-psa.R"))
source(here::here("functions", "function_plot-psa.R"))

``` 

Set Parameters with static values

```{r}
### Hyperparameters
cycle_length <- 1/12                              # cycle length equals 1 month
n.t   <- 12                                       # number of cycles, time horizon equals 1 year 
d.c   <- d.e <- 0.05 / (1/cycle_length)           # discounting of costs and QALYs, annual rate = 5%
v.n   <- c("NE", "VR", "MB", "NMB", "OT")         # Model states  
v.M_1 <- c(1, 0, 0, 0, 0)                         # Everyone begins in no event state

#Number of MC Samples
n_samples <- 1000
#Set the random number seed for reproducibility
set.seed(123) 

```

Set MC draws for event rate parameters

```{r}
##Rivaroxaban

#Health States 
#Event rates per person-6months converted to person-year:
sample_r.ne_vr   <- runif(n_samples, min = 0, max = 0.000 + 0.50) * 2   # rate of no event to VTE recurrence 
sample_r.ne_mb   <- runif(n_samples, min = 0, max = 0.019 + 0.05) * 2   # rate of no event to  major bleed 
sample_r.ne_nmb  <- runif(n_samples, min = 0, max = 0.125 + 0.05) * 2   # rate of no event to CRNMB
sample_r.nmb_ot  <- runif(n_samples, min = 0, max = 0.20) * 2   # rate of CNMB to come off treatment  
sample_r.mb_ot   <- runif(n_samples, min = 0, max = 0.20) * 2   # rate of MB to come off treatment 

#Transition probabilities (per month cycle): 
sample_pr.ne_vr   <- 1 - exp(-sample_r.ne_vr * cycle_length)        # from no event to VTE recurrence 
sample_pr.ne_mb   <- 1 - exp(-sample_r.ne_mb * cycle_length)        # from no event to  major bleed 
sample_pr.ne_nmb  <- 1 - exp(-sample_r.ne_nmb * cycle_length)       # from no event to CRNMB
sample_pr.nmb_ot  <- 1 - exp(-sample_r.nmb_ot * cycle_length)       # from CNMB to come off treatment  
sample_pr.mb_ot   <- 1 - exp(-sample_r.mb_ot * cycle_length)        # from MB to come off treatment  

# Create dataframe of lists of parameter values for state transition probabilities
l.probr <- list()
for (i in 1:n_samples){
  sample.l.probr <- list(pr.ne_vr = sample_pr.ne_vr[i],
                         pr.ne_mb = sample_pr.ne_mb[i],
                         pr.ne_nmb = sample_pr.ne_nmb[i],
                         pr.ne_ne = 1 - sum(sample_pr.ne_vr[i],
                                            sample_pr.ne_mb[i],
                                            sample_pr.ne_nmb[i]),
                         pr.nmb_ot = sample_pr.nmb_ot[i],
                         pr.mb_ot = sample_pr.mb_ot[i])
  l.probr[i] <- list(sample.l.probr)
}

df.probr_riv <- as.data.frame(do.call(rbind, l.probr))

```

```{r}
##Warfarin

#Health States 
#Event rates per person-6months converted to person-year:
sample_r.ne_vr   <- runif(n_samples, min = 0, max = 0.026 + 0.50) * 2   # rate of no event to VTE recurrence 
sample_r.ne_mb   <- runif(n_samples, min = 0, max = 0.090 + 0.05) * 2   # rate of no event to  major bleed 
sample_r.ne_nmb  <- runif(n_samples, min = 0, max = 0.308 + 0.05) * 2   # rate of no event to CRNMB
sample_r.nmb_ot  <- runif(n_samples, min = 0, max = 0.20) * 2   # rate of CNMB to come off treatment  
sample_r.mb_ot   <- runif(n_samples, min = 0, max = 0.20) * 2   # rate of MB to come off treatment 

#Transition probabilities (per month cycle): 
sample_pr.ne_vr   <- 1 - exp(-sample_r.ne_vr * cycle_length)        # from no event to VTE recurrence 
sample_pr.ne_mb   <- 1 - exp(-sample_r.ne_mb * cycle_length)        # from no event to  major bleed 
sample_pr.ne_nmb  <- 1 - exp(-sample_r.ne_nmb * cycle_length)       # from no event to CRNMB
sample_pr.nmb_ot  <- 1 - exp(-sample_r.nmb_ot * cycle_length)       # from CNMB to come off treatment  
sample_pr.mb_ot   <- 1 - exp(-sample_r.mb_ot * cycle_length)        # from MB to come off treatment  

# Create dataframe of lists of parameter values for state transition probabilities
l.probr <- list()
for (i in 1:n_samples){
  sample.l.probr <- list(pr.ne_vr = sample_pr.ne_vr[i],
                         pr.ne_mb = sample_pr.ne_mb[i],
                         pr.ne_nmb = sample_pr.ne_nmb[i],
                         pr.ne_ne = 1 - sum(sample_pr.ne_vr[i],
                                            sample_pr.ne_mb[i],
                                            sample_pr.ne_nmb[i]),
                         pr.nmb_ot = sample_pr.nmb_ot[i],
                         pr.mb_ot = sample_pr.mb_ot[i])
  l.probr[i] <- list(sample.l.probr)
}

df.probr_war <- as.data.frame(do.call(rbind, l.probr))

```

Set MC draws for cost parameters: gamma distributions

```{r}
#' @param v.cost is a numeric vector containing cost values for each state

##Rivaroxaban Costs (per 6 months (year/2))
riv.c.ne      <- rgamma(n_samples, shape=10, rate=1/14) / 12
              #Cost for staying healthy in rivaroxaban arm
riv.c.vr      <- rgamma(n_samples, shape=10, rate=1/21.5)  / 12
              #Cost for VTE recurrence in Rivaroxaban arm ($215 annually) 
riv.c.mb      <- rgamma(n_samples, shape=2, rate= 1/245)  / 12
              #Cost for major bleed in Rivaroxaban arm ($490 annually)
riv.c.nmb     <- rgamma(n_samples, shape=2, rate= 1/152)  / 12
              #Cost for CRNMB in Rivaroxaban arm ($304 annually)
riv.c.ot      <- rgamma(n_samples, shape=1, rate=1/107.5) / 12 #Cost of off treatment ($0 annually)

##Warfarin Costs (per 6 months (year/2))
war.c.ne      <- rgamma(n_samples, shape=10, rate=1/16.5)  / 12
              #Cost for staying healthy in Warfarin arm ($165 annually)
war.c.vr      <- rgamma(n_samples, shape=10, rate=1/26.7)  / 12
              #Cost for VTE recurrence in Warfarin arm ($267 annually)  
war.c.mb      <- rgamma(n_samples, shape=3, rate= 1/217.5)  / 12
              #Cost for major bleed in Warfarin arm ($435 annually)
war.c.nmb     <- rgamma(n_samples, shape=3, rate= 1/166) / 12
              #Cost for CRNMB in Warfarin arm ($332 annually)
war.c.ot      <- rgamma(n_samples, shape=1, rate=1/133.5) / 12  #Cost of off treatment ($0 annually)

##Vectors for cost values
riv.v.cost <- cbind(riv.c.ne, riv.c.vr, riv.c.mb, riv.c.nmb, riv.c.ot)
war.v.cost <- cbind(war.c.ne, war.c.vr, war.c.mb, war.c.nmb, war.c.ot)

l.costr <- list()
l.costw <- list()
for (i in 1:n_samples){
  l.costr[[i]] = unlist(riv.v.cost[i, ], use.names=F)
  l.costw[[i]] = unlist(war.v.cost[i, ], use.names=F)
}

```

Set MC draws for utility parameters: logit distributions

```{r}
#' @param v.util is a numeric vector containing utility values for each state

##Rivaroxaban Values 
riv.u.ne    <- rlogis(n_samples, location=0.825, scale=0.02)        / 12  #Utility of No Event
riv.u.vr    <- rlogis(n_samples, location=0.76, scale=0.063774925)  / 12  #Utility of recurrence  
riv.u.mb    <- rlogis(n_samples, location=0.61, scale=0.274593292)  / 12  #Utility of major bleed
riv.u.nmb   <- rlogis(n_samples, location=0.65, scale=0.115559069)  / 12  #Utility of CRNMB 
riv.u.ot    <- rlogis(n_samples, location=0.68, scale=0.063774925)  / 12  #Utility of off treatment 
##Warfarin Values 
war.u.ne    <- rlogis(n_samples, location=0.825, scale=0.02)        / 12  #Utility of No Event
war.u.vr    <- rlogis(n_samples, location=0.76, scale=0.063774925)  / 12  #Utility of recurrence  
war.u.mb    <- rlogis(n_samples, location=0.61, scale=0.274593292)  / 12  #Utility of major bleed
war.u.nmb   <- rlogis(n_samples, location=0.65, scale=0.115559069)  / 12  #Utility of CRNMB 
war.u.ot    <- rlogis(n_samples, location=0.68, scale=0.063774925)  / 12  #Utility of off treatment 

##Vectors for utility values
riv.v.util <- cbind(riv.u.ne, riv.u.vr, riv.u.mb, riv.u.nmb, riv.u.ot)
war.v.util <- cbind(war.u.ne, war.u.vr, war.u.mb, war.u.nmb, war.u.ot)

l.utilr <- list()
l.utilw <- list()
for (i in 1:n_samples){
  l.utilr[[i]] = unlist(riv.v.util[i, ], use.names=F)
  l.utilw[[i]] = unlist(war.v.util[i, ], use.names=F)
}

```

Generate params

```{r}

## Generate an input dataframe containing parameters for each simulation
riv.input_params <- as.data.frame(cbind(nsim = as.character(seq(1, n_samples)),
                                        n.t = as.numeric(rep(n.t, n_samples)),
                                        d.c = as.numeric(rep(d.c, n_samples)),
                                        d.e = as.numeric(rep(d.e, n_samples)),
                                        v.n = rep(list(v.n), n_samples),
                                        v.cost = l.costr,  ### SENS
                                        v.util = l.utilr,  ### SENS
                                        v.M_1 = rep(list(v.M_1), n_samples),
                                        pr.ne_ne = df.probr_riv$pr.ne_ne,
                                        pr.ne_vr = df.probr_riv$pr.ne_vr,
                                        pr.ne_mb = df.probr_riv$pr.ne_mb,
                                        pr.ne_nmb = df.probr_riv$pr.ne_nmb,
                                        pr.mb_ot = df.probr_riv$pr.mb_ot,
                                        pr.nmb_ot = df.probr_riv$pr.nmb_ot
                                        ))

war.input_params <- as.data.frame(cbind(nsim = as.character(seq(1, n_samples)),
                                        n.t = as.numeric(rep(n.t, n_samples)),
                                        d.c = as.numeric(rep(d.c, n_samples)),
                                        d.e = as.numeric(rep(d.e, n_samples)),
                                        v.n = rep(list(v.n), n_samples),
                                        v.cost = l.costw,  ### SENS
                                        v.util = l.utilw,  ### SENS
                                        v.M_1 = rep(list(v.M_1), n_samples),
                                        pr.ne_ne = df.probr_war$pr.ne_ne,
                                        pr.ne_vr = df.probr_war$pr.ne_vr,
                                        pr.ne_mb = df.probr_war$pr.ne_mb,
                                        pr.ne_nmb = df.probr_war$pr.ne_nmb,
                                        pr.mb_ot = df.probr_war$pr.mb_ot,
                                        pr.nmb_ot = df.probr_war$pr.nmb_ot
                                        ))

```

Run Markov models on sensitivity values

```{r}

### Riv
#Initialize dataframe to store results
sim_results_riv <- data.frame(nsim = as.character(seq(1:n_samples)))
  
for (i in 1:n_samples){
  sim_markov_riv <- Markov(n.t = riv.input_params$n.t[[i]], 
                           d.c = riv.input_params$d.c[[i]],  
                           d.e = riv.input_params$d.e[[i]],  
                           v.n = riv.input_params$v.n[[i]],  
                           v.cost = riv.input_params$v.cost[[i]],  
                           v.util = riv.input_params$v.util[[i]],  
                           v.M_1 = riv.input_params$v.M_1[[i]],  
                           pr.ne_ne = riv.input_params$pr.ne_ne[[i]],  
                           pr.ne_vr = riv.input_params$pr.ne_vr[[i]],  
                           pr.ne_mb = riv.input_params$pr.ne_mb[[i]],  
                           pr.ne_nmb = riv.input_params$pr.ne_nmb[[i]],  
                           pr.mb_ot = riv.input_params$pr.mb_ot[[i]],  
                           pr.nmb_ot = riv.input_params$pr.nmb_ot[[i]])
  sim_results_riv$trace[i] <- list(sim_markov_riv$m.TRr)
  sim_results_riv$tc_hat[i] <- sim_markov_riv$tc_hat
  sim_results_riv$te_hat[i] <- sim_markov_riv$te_hat
}

### War
#Initialize dataframe to store results
sim_results_war <- data.frame(nsim = as.character(seq(1:n_samples)))
  
for (i in 1:n_samples){
  sim_markov_war <- Markov(n.t = war.input_params$n.t[[i]], 
                           d.c = war.input_params$d.c[[i]],  
                           d.e = war.input_params$d.e[[i]],  
                           v.n = war.input_params$v.n[[i]],  
                           v.cost = war.input_params$v.cost[[i]],  
                           v.util = war.input_params$v.util[[i]],  
                           v.M_1 = war.input_params$v.M_1[[i]],  
                           pr.ne_ne = war.input_params$pr.ne_ne[[i]],  
                           pr.ne_vr = war.input_params$pr.ne_vr[[i]],  
                           pr.ne_mb = war.input_params$pr.ne_mb[[i]],  
                           pr.ne_nmb = war.input_params$pr.ne_nmb[[i]],  
                           pr.mb_ot = war.input_params$pr.mb_ot[[i]],  
                           pr.nmb_ot = war.input_params$pr.nmb_ot[[i]])
  sim_results_war$trace[i] <- list(sim_markov_war$m.TRr)
  sim_results_war$tc_hat[i] <- sim_markov_war$tc_hat
  sim_results_war$te_hat[i] <- sim_markov_war$te_hat
}

```

Calculate ICERs

```{r}

sim_results_cost <- data.frame(war = sim_results_war$tc_hat,
                               riv = sim_results_riv$tc_hat)
sim_results_effectiveness <- data.frame(war = sim_results_war$te_hat,
                                        riv = sim_results_riv$te_hat)

psa <- make_psa_obj(cost = sim_results_cost, effectiveness = sim_results_effectiveness, parameters = NULL,
                    strategies = c("Warfarin", "Rivaroxaban"), currency = "$", other_outcome = NULL)

# Save psa object
saveRDS(psa, here::here("output","results_1yr_sens_08_MC_CostsUtility.rds"))

# See mean ICER
summary.psa(psa, calc_sds=TRUE)

```

Distribution of ICER estimates

Table
```{r}
sim_results_ICER <- as.data.frame(
  cbind(inc_cost = sim_results_cost$riv - 
                   sim_results_cost$war,
        inc_effectiveness = sim_results_effectiveness$riv -                                           sim_results_effectiveness$war)) %>%
  mutate(ICER = inc_cost/inc_effectiveness,
         name = "MC")

mean(sim_results_ICER$ICER)
sd(sim_results_ICER$ICER)

library(boot)
set.seed(2025)
boot.icer = boot(data=sim_results_ICER$ICER,
                 R=1000,
                 statistic=function(x,id){ mean(x[id])})
boot.ci(boot.icer, type="bca")

```

Plot (ICERs)

```{r}
basecase <- readRDS(here::here("output","results_1yr_basecase.rds"))

sim_results_ICER_boxplot <- sim_results_ICER %>%
  add_row(inc_cost = basecase$delta.c,
          inc_effectiveness = basecase$delta.e,
          ICER = basecase$ICER,
          name = "base")

ggplot(sim_results_ICER_boxplot, aes(x=name, y=ICER)) + 
  geom_boxplot() + 
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = c(-5000, 0)) +
  coord_flip() +
  xlab("Analysis") + 
  ylab("ICER ($/QALY)")

```

PSA plot (CERs)

```{r}

plot.psa(psa) +
  ggthemes::scale_color_colorblind() +
  ggthemes::scale_fill_colorblind() +
  xlab("Effectiveness (QALYs)") +
  guides(col = guide_legend(nrow = 2)) +
  theme(legend.position = "right")

```

Range of CERs (highest cost / lowest effectiveness)

```{r}
data.frame(
  Strategy = c("Warfarin", "Rivaroxaban"),
  max_cost = c(max(psa$cost$Warfarin), max(psa$cost$Rivaroxaban)),
  min_effectiveness = c(min(psa$effectiveness$Warfarin), min(psa$effectiveness$Rivaroxaban)),
  worst_CER = c(max(psa$cost$Warfarin)/min(psa$effectiveness$Warfarin),
                max(psa$cost$Rivaroxaban)/min(psa$effectiveness$Rivaroxaban))
)
```

% of CER simulations where rivaroxaban is more cost-effective than warfarin

```{r}
# Simulated CERs for warfarin
psa_cer_war <- round(psa$cost$Warfarin / psa$effectiveness$Warfarin, digits=2)
# Simulated CERs for rivaroxaban
psa_cer_riv <- round(psa$cost$Rivaroxaban / psa$effectiveness$Rivaroxaban, digits=2)

# simulated CERs for rivaroxaban that are more cost effective than warfarin
psa_cer <- expand.grid(war = psa_cer_war, riv = psa_cer_riv) %>%
  distinct() %>%
  mutate(riv_costeffective = ifelse(riv < war, 1, 0))

sum(psa_cer$riv_costeffective)/nrow(psa_cer)

```

% of CER simulations where rivaroxaban is more cost-effective than the best warfarin CER

```{r}
# Simulated CERs for warfarin
psa_cer_war <- round(psa$cost$Warfarin / psa$effectiveness$Warfarin, digits=2)
psa_cer_war <- sort(psa_cer_war)
# Simulated CERs for rivaroxaban
psa_cer_riv <- round(psa$cost$Rivaroxaban / psa$effectiveness$Rivaroxaban, digits=2)
psa_cer_riv <- sort(psa_cer_riv)

# simulated CERs for rivaroxaban that are more cost effective than best warfarin
# i.e., lower than the minimum CER for warfarin
psa_cer_riv_costeffective <- psa_cer_riv[psa_cer_riv < psa_cer_war[1]]
length(psa_cer_riv_costeffective) / length(psa_cer_riv)
```